Measurement of total hadronic differential cross sections in the LArIAT experiment

### 2018-05-21 Chapter-5/5-Preparatory.tex
s

\section{Estimate of Backgrounds in the Pion Cross Section}\label{ch:PionXSBkgSub}
We use the beamline simulation and the DDMC simulation to estimate the background in the total hadronic pion cross section. Two categories of background exists for the negative pion cross section measurement: the one related to the pion interaction in the chamber, discussed in Section \ref{ch:CaptureAndDecay} and the one related to the beamline contamination, discussed in Section \ref{ch:PionXSBkgSub2}.

@@

\subsection{Beamline Background}\label{ch:PionXSBkgSub2}
As shown in Section \ref{ch:beamlineComposition}, we use G4Beamline to estimate the beam composition at WC4, obtaining the composition shown in Table \ref{tab:beamline}.
Once we have estimated the beam composition at WC4, the next step to estimate the beamline background in the cross section measurement by propagating pions, muons and electrons to the TPC and evaluate their contribution to the cross section. To do so, we simulate the same number of electrons, muons and pions with the DDMC and we apply the same selection filters on the three samples. The number of events per particle species surviving this selection is shown on table \ref{tab:MCafterCutContaminants}.

@@

\begin{figure}[p]
\centering
%\includegraphics[width=\textwidth]{Studies/Figures/Staggered60A.png}
\caption{Left: staggered contributions to the interacting kinetic energy distribution for electron (grey), muons (yellow) and pion (blue) in the 60A simulation sample. Right: staggered contributions to the incident kinetic energy distribution for electron (grey), muons (yellow) and pion (blue) in the 60A simulation sample.  }
\label{fig:stag60A}
\end{figure}

>>>>

\section{Estimate of Backgrounds in the Pion Cross Section}\label{ch:PionXSBkgSub}

We use the beamline simulation and the DDMC simulation to estimate the background in the total hadronic pion cross section. Two categories of background exists for the negative pion cross section measurement: the one related to the pion interaction in the chamber, discussed in Section \ref{ch:CaptureAndDecay} and the one related to the beamline contamination, discussed in Section \ref{ch:PionXSBkgSub2}.

@@

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Beamline Background}\label{ch:PionXSBkgSub2}
We define beamline background every TPC track matched to the WC track which is not a primary pion. Potentially, there are 4 different types of beamline background:
\begin{itemize}
\item[]1) electrons,
\item[]2) muons,
\item[]3) secondaries from pion events,
\item[]4) matched pile up events.
\end{itemize}

The first step is to estimate what percentage of events used in the cross section calculation is not a primary pion.  We start by noting that the last type of background, the ``matched pile up" events, is a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match with a halo particle in the absence of a beamline particle\footnote{ Events with multiple WC2TPC matches are always rejected.} negligibly small. %\textcolor{red}{SHOW VTX distribution in WC2TPC match}

As shown in Section \ref{ch:beamlineComposition}, we use G4Beamline to estimate the beam composition at WC4, obtaining the composition shown in Table \ref{tab:beamline}.
The next step to estimate the beamline background in the cross section measurement is propagating pions, muons and electrons to the TPC and evaluate their contribution to the cross section. To do so, we simulate the same number of electrons, muons and pions with the DDMC and we apply the same selection filters on the three samples. The number of events per particle species surviving this selection is shown on table \ref{tab:MCafterCutContaminants}.

Pions can travel the length of the LArIAT beamline and interact hadronically in the steel or in the non-instrumented argon upstream to the TPC front face. Or, they could decay in flight between WC4 and the TPC. One of the interaction products can leak into the TPC and be matched with the WC track, contributing to the pool of events used for the cross section calculation. We call this type of particles ``secondaries" from pion events, with a terminology inspired by Geant4. 
We estimate the number of secondaries using the DDMC pion sample.  The percentage of secondaries is given by the number of matched WC2TPC tracks whose corresponding particle is not flagged as primary by Geant4.  The secondary to pion ratio is 4.9\% in the 60A sample and 4.3\% in the 100A sample.

@@

\begin{figure}[p]
\centering
\includegraphics[width=\textwidth]{Chapter-5/Images/Background60A.pdf}
\caption{Left: staggered contributions to the interacting kinetic energy distribution for electron (grey), muons (yellow) and pion (blue) in the 60A simulation sample. Right: staggered contributions to the incident kinetic energy distribution for electron (grey), muons (yellow) and pion (blue) in the 60A simulation sample.  }
\label{fig:stag60A}
\end{figure}

### 2018-05-21 Chapter-6/6-PionCrossSection.tex
s

\subsection{Background subtraction}\label{ch:BKGsubXS}
Even if pions are by far the biggest component of the beam in negative polarity runs, the LArIAT beam is not a pure pion beam. While useful to discriminate  pions/muons/electrons from kaons, and protons, the beamline detectors are not sensitive enough to  discriminate among the lighter particles in the beam: electrons, muons and pions fall under the same mass hypothesis. Thus, we need to assess the background from beamline particles other than pions in the event selections used for the pion cross section analysis and correct for its effects.

\subsubsection{Beam Composition}\label{sec:BeamAtWC4}
We define beamline background every TPC track matched to the WC track which is not a primary pion. Potentially, there are 4 different types of beamline background:
\begin{itemize}
\item[]1) electrons,
\item[]2) muons,
\item[]3) secondaries from pion events,
\item[]4) matched pile up events.
\end{itemize}

The first step is to estimate what percentage of events used in the cross section calculation is not a primary pion.  The next two sections will illustrate this estimate for the electrons, muons and secondaries from pion event.
We estimate the last type of background, the ``matched pile up" events, to be a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match with a halo particle in the absence of a beamline particle\footnote{ Events with multiple WC2TPC matches are always rejected.} negligibly small. \textcolor{red}{SHOW VTX distribution in WC2TPC match}

\subsubsection{Background from Beamline Electrons and Muons}\label{stionEMu}
\begin{figure}[b]
%\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Studies/Figures//Beam60A.png}
%\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Studies/Figures//Beam100A.png}
\caption{Beam composition for the -60A runs (left) and -100A runs (right). The solid blue plot represents the simulated pion content, the yellow plot represents the simulated muon content and the grey plot represents the simulated electron content. The plots are area normalized to the number of data events, shown in red. }
\label{fig:BeamComposition}
\end{figure}

We estimate the percentage of electrons and muons in the beam via the G4Beamline MC. 
Since the beamline composition is a function of the magnet settings, we simulate separately events for magnet current of -60A and -100A. 
Table \ref{tab:beamline} shows the beam composition per magnet setting after the mass selection according to the G4Beamline simulation.
\begin{table}[p]
\centering
\begin{tabular}{|l|c|c|}
\hline
                     & I = -60 A           & I = -100 A \\ \hline
G4Pions       &   68.8 \%           &      87.4 \%        \\ \hline
G4Muons     &     4.6 \%           &        3.7 \%         \\ \hline
G4Electrons &   26.6 \%           &        8.9 \%        \\ \hline
\end{tabular}
\caption{Simulated beamline composition per magnet settings}
\label{tab:beamline}
\end{table}

Figure \ref{fig:BeamComposition} shows the momentum predictions from G4Beamline overlaid with data for the 60A runs (left) and for the 100A runs (right). The predictions for electrons, muons and pions have been staggered and their sum is area normalized to data, which is shown in red. Albeit not perfect, these plots show a reasonable agreement between the momentum shapes in data and MC. We attribute  the difference in shape to the lack of simulation of the WC efficiency in the MC which is momentum dependent and leads to enhance the number events in the center of the momentum distribution.

Once the beam composition at WC4 is know,  we simulate the electrons, muons and pions with the DDMC and we subject the three samples to the same selection chain (WC2TPC match, shower filter, pile up filter). The percentage of electrons and muons surviving the selection chain weighted by the beam composition is the  electron and muon background in the pion cross section sample, as shown in Table \ref{tab:MCafterCutContaminants}.

\subsubsection{Background from secondaries at TPC Front Face}
Pions can travel the length of the LArIAT beamline and interact hadronically in the steel or in the non-instrumented argon upstream to the TPC front face. Or, they could decay in flight between WC4 and the TPC. One of the interaction products can leak into the TPC and be matched with the WC track, contributing to the pool of events used for the cross section calculation. We call this type of particles ``secondaries" from pion events, with a terminology inspired by Geant4. 
We estimate the number of secondaries using the DDMC pion sample.  The percentage of secondaries is given by the number of matched WC2TPC tracks whose corresponding particle is not flagged as primary by Geant4.  The secondary to pion ratio is 4.9\% in the 60A sample and $Y$\% in the 100A sample.

>>>>

\subsection{Background subtraction}\label{ch:BKGsubXS}
%Even if pions are by far the biggest component of the beam in negative polarity runs, the LArIAT beam is not a pure pion beam. While useful to discriminate  pions/muons/electrons from kaons, and protons, the beamline detectors are not sensitive enough to  discriminate among the lighter particles in the beam: electrons, muons and pions fall under the same mass hypothesis. Thus, we need to assess the background from beamline particles other than pions in the event selections used for the pion cross section analysis and correct for its effects.

%\subsubsection{Beam Composition}\label{sec:BeamAtWC4}
%We define beamline background every TPC track matched to the WC track which is not a primary pion. Potentially, there are 4 different types of beamline background:
%\begin{itemize}
%\item[]1) electrons,
%\item[]2) muons,
%\item[]3) secondaries from pion events,
%\item[]4) matched pile up events.
%\end{itemize}

%The first step is to estimate what percentage of events used in the cross section calculation is not a primary pion.  The next two sections will illustrate this estimate for the electrons, muons and secondaries from pion event.
%We estimate the last type of background, the ``matched pile up" events, to be a negligible fraction, because of the definition of the WC2TPC match: we deem the probability of a single match with a halo particle in the absence of a beamline particle\footnote{ Events with multiple WC2TPC matches are always rejected.} negligibly small. \textcolor{red}{SHOW VTX distribution in WC2TPC match}

%\subsubsection{Background from Beamline Electrons and Muons}\label{stionEMu}
%\begin{figure}[b]
%\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Studies/Figures//Beam60A.png}
%\includegraphics[width=0.5\textwidth,height=\textheight,keepaspectratio]{Studies/Figures//Beam100A.png}
%\caption{Beam composition for the -60A runs (left) and -100A runs (right). The solid blue plot represents the simulated pion content, the yellow plot represents the simulated muon content and the grey plot represents the simulated electron content. The plots are area normalized to the number of data events, shown in red. }
%\label{fig:BeamComposition}
%\end{figure}

%We estimate the percentage of electrons and muons in the beam via the G4Beamline MC. 
%Since the beamline composition is a function of the magnet settings, we simulate separately events for magnet current of -60A and -100A. 
%Table \ref{tab:beamline} shows the beam composition per magnet setting after the mass selection according to the G4Beamline simulation.
%\begin{table}[p]
%\centering
%\begin{tabular}{|l|c|c|}
%\hline
 %                   & I = -60 A           & I = -100 A \\ \hline
%G4Pions       &   68.8 \%           &      87.4 \%        \\ \hline
%G4Muons     &     4.6 \%           &        3.7 \%         \\ \hline
%G4Electrons &   26.6 \%           &        8.9 \%        \\ \hline
%\end{tabular}
%\caption{Simulated beamline composition per magnet settings}
%\label{tab:beamline}
%\end{table}

%Figure \ref{fig:BeamComposition} shows the momentum predictions from G4Beamline overlaid with data for the 60A runs (left) and for the 100A runs (right). The predictions for electrons, muons and pions have been staggered and their sum is area normalized to data, which is shown in red. Albeit not perfect, these plots show a reasonable agreement between the momentum shapes in data and MC. We attribute  the difference in shape to the lack of simulation of the WC efficiency in the MC which is momentum dependent and leads to enhance the number events in the center of the momentum distribution.

%Once the beam composition at WC4 is know,  we simulate the electrons, muons and pions with the DDMC and we subject the three samples to the same selection chain (WC2TPC match, shower filter, pile up filter). The percentage of electrons and muons surviving the selection chain weighted by the beam composition is the  electron and muon background in the pion cross section sample, as shown in Table \ref{tab:MCafterCutContaminants}.

%\subsubsection{Background from secondaries at TPC Front Face}
%Pions can travel the length of the LArIAT beamline and interact hadronically in the steel or in the non-instrumented argon upstream to the TPC front face. Or, they could decay in flight between WC4 and the TPC. One of the interaction products can leak into the TPC and be matched with the WC track, contributing to the pool of events used for the cross section calculation. We call this type of particles ``secondaries" from pion events, with a terminology inspired by Geant4. 
We estimate the number of secondaries using the DDMC pion sample.  The percentage of secondaries is given by the number of matched WC2TPC tracks whose corresponding particle is not flagged as primary by Geant4.  The secondary to pion ratio is 4.9\% in the 60A sample and $Y$\% in the 100A sample.
